---
title: "Figures "
output: html_notebook
---


1) bubble plots pathways
2) barplots types omics per community
### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/3_layers_integration/")) 
```

```{r}
library(ggplot2)
library(dplyr)
```
## load data




```{r}
table_nodes <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/data/network/Table_nodesAA_3_layers_filt_consensus.csv")

#table_nodes <- table_nodes[table_nodes$Community %in% c(1:7),]

results_gsea <- read.csv("results/network/Results_centrality_networksAA_3_layers_filt_consensus.csv")
```

```{r}
table_edges <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/network/network_correlation_positive_AA_3_layers_filt_consensus.csv")
```


```{r}
features_mofa <- read.csv("processing/extracted_MOFA_features.csv")
features_mofa <- features_mofa[,1:9]
features_mofa$item[grepl("Met", features_mofa$item)] <- features_mofa$met[grepl("Met", features_mofa$item)]
names(features_mofa)[2] <- "feat1"
features_mofa$MOFA <- "yes"
```

```{r}
mdm <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/processing_microbione_derived_metabolomics.csv")
names(mdm)[2] <- "feat1"
```

```{r}
names(table_nodes)
table_nodes <- merge(table_nodes, features_mofa, by = "feat1", all.x = TRUE)
table(table_nodes$MOFA)
```
```{r}
table_edges_2 <- table_edges[table_edges$feat1 %in% mdm$feat1 | table_edges$feat2 %in% mdm$feat1, ]
table_edges_2_features <- unique(c(table_edges_2$feat1, table_edges_2$feat2))
table(table_nodes$MOFA)
```


```{r}
table_nodes$MDM <- ifelse(table_nodes$feat1 %in% table_edges_2_features, "yes", "no")
```

```{r}
write.csv(table_nodes, "results/network/table_nodes_consensus_network_with_MDM_and_MOFA.csv")
```

```{r}
col <- c("#CC6600", "#009933", "#0066CC","#CC00CC","#009999","#990033")
```

```{r}
table(table_nodes$Community)
```

```{r}
results_gsea$Terms_Lipids <- as.vector(results_gsea$Terms_Lipids)
results_gsea$Terms_Lipids <- gsub(",", "\n", results_gsea$Terms_Lipids)
results_gsea$Terms_Lipids[results_gsea$Terms_Lipids == "Not"] <- c("No lipids in this community_", "No lipids in this community.","No lipids in this community")
results_gsea$Terms_Lipids[results_gsea$Terms_Lipids == "triacylglycerol\ndiacylglycerol"] <- c("triacylglycerol\ndiacylglycerol", "triacylglycerol\ndiacylglycerol_")
```


```{r}
table <- results_gsea
table$value <- 50
table <- table[order(table$Number_community, decreasing = TRUE),]
table$color <- col
position <- table$Terms_Lipids
```

```{r}
cairo_pdf("results/figures/bubble_plot_pathways_lipidomics.pdf", height = 3, width = 7,family="mono")

a <- ggplot(table, aes(x = value, y = Terms_Lipids, size = N_metabolites, color = as.factor(Number_community)))+ geom_point()+
  labs(size ="Number of features")
a + theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank(), axis.text.y = element_text(size=12, color = "black"))+ coord_cartesian(xlim = c(49, 50.05), expand = TRUE) + scale_y_discrete(limits = position, position = "right")+ 
    scale_size_continuous(range = c(3, 8))+ scale_color_manual(values= col)+ guides(color = FALSE)

dev.off()

```

```{r}
results_gsea$Terms_Metabolites <- as.vector(results_gsea$Terms_Metabolites)
results_gsea$Terms_Metabolites[results_gsea$Terms_Metabolites == "Not"] <- "No metabolites in this community"
results_gsea$Terms_Metabolites[results_gsea$Terms_Metabolites == ""] <- c("No pathway significant_", "No pathway significant")
```


```{r}
table <- results_gsea
table$value <- 50
table <- table[order(table$Number_community, decreasing = TRUE),]
table$color <- col
position <- table$Terms_Metabolites
```

```{r}
cairo_pdf("results/figures/bubble_plot_pathways_metabolomics.pdf", height = 2, width = 7,family="mono")

a <- ggplot(table, aes(x = value, y = Terms_Metabolites, size = N_metabolites, color = as.factor(table$Number_community)))+ geom_point()+
  labs(size ="Number of features")
a + theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank(), axis.text.y = element_text(size=12, color = "black"))+ coord_cartesian(xlim = c(49, 50.05), expand = TRUE) + scale_y_discrete(limits = position, position = "right")+ 
    scale_size_continuous(range = c(3, 8))+ scale_color_manual(values= col)+ guides(color = FALSE)

dev.off()
```

```{r}
met <- read.csv("/home/flomik/Desktop/Code-PHD/COCOMO_v2/processing/metabolites_pathways.csv")
lip <- read.csv("/home/flomik/Desktop/Code-PHD/COCOMO_v2/processing/lipids_pathways.csv")
names(lip) <- names(met)
annot <- rbind(met, lip)
```

```{r}
names(table_nodes)[4] <- "BIOCHEMICAL"
table_nodes_2 <- merge(table_nodes, annot, by = "BIOCHEMICAL", all.x = TRUE)
table_nodes_2$type_2 <- paste0(table_nodes_2$type_1, "_", table_nodes_2$SUPER.PATHWAY)

```

```{r}
table_communities <- data.frame(table(table_nodes$type_1, table_nodes$Community))
table_communities_2 <- data.frame(table(table_nodes_2$type_2, table_nodes_2$Community))
table_communities_2 <- table_communities_2[table_communities_2$Freq > 0, ]
```

```{r}
ggplot(table_communities, aes(fill=Var1, y=Freq, x=Var2, label = Freq)) + 
    geom_bar(stat="identity", alpha = 0.7,  colour="black")+
  geom_text(size = 4, position = position_stack(vjust = 0.5))+
  theme(
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+theme_minimal()

ggsave(paste0("results/figures/barplot_type_network_whole.pdf"), height = 6, width = 6)
```

ggplot(d3, aes(fill=Category, y=fraction, x=Group))+ 
    geom_bar(stat="identity", alpha = 1, width = .7, colour="white", lwd=0.1) +
  labs(x="Comparison",y="Comorbidities frequency")+
    scale_fill_manual(values = rev(mycols))+
  geom_text(aes(label=ifelse(fraction >= 0.07, paste0("n = ", count, ", \n", round(fraction*100, 0), "%"),"")), position=position_stack(vjust=0.5), colour="white")+
  coord_flip() +
  scale_y_continuous(labels = percent_format()) +
  labs(y="", x="")

ggsave(paste0("results/figures/barplot_comparisons_LIMMA_COCOMO.pdf"), width = 15, height = 8)

```{r}
col_1 <- colorRampPalette(c("#2f2e75", "#dbd9fc"))
col_2 <- colorRampPalette(c("#6b0c53", "#fcd9f0"))
col3 <- col_2(8) 

col <- c(col_1(10),col3[1:4],"#4eaebf",col3[4:7],"#5f946f")
```

```{r}
data_new2 <- table_communities_2 %>%                                      # Top N highest values by group
  arrange(desc(Freq)) %>% 
  group_by(Var2) %>%
  slice(1:3)
```
```{r}
m <- seq(1, 27, by = 3)
n <- c(1, 4, 7, 10, 13, 16, 19, 20, 23)
```


```{r}
list_lab <- list()
y = 1
for (i in n) {
  list_lab[[y]] <- paste0("\n", data_new2$Var1[i], "\n", data_new2$Var1[i+1], "\n", data_new2$Var1[i+2])
  y = y +1
  
  
}

list_lab[[7]] <- gsub("NA.*", "NA", list_lab[[7]])
```

```{r}
list_lab_2 <- c(rep(list_lab[[1]], 8),
                rep(list_lab[[2]], 9),
                rep(list_lab[[3]], 6),
                rep(list_lab[[4]], 6),
                rep(list_lab[[5]], 10),
                rep(list_lab[[6]], 3),
                rep(list_lab[[7]], 1),
                rep(list_lab[[8]], 3),
                rep(list_lab[[9]], 4))

table_communities_2$Var2 <- paste0(table_communities_2$Var2, list_lab_2)
```

```{r}
ggplot(table_communities_2, aes(fill=Var1, y=Freq, x=Var2, label = Freq, color = Var1)) + 
    geom_bar(stat="identity", alpha = 0.7,  colour="black")+
  theme(
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  geom_text(aes(label=ifelse(Freq >= 20, Freq, "")), size = 4, position = position_stack(vjust = 0.5), color = "black")+scale_fill_manual(values = col) + scale_x_discrete(position = "top") 

ggsave(paste0("results/network/barplot_type_network_communties_3.pdf"), height = 6, width = 19)
```


```{r}
i <- 1
```

```{r}
table_nodes <- table_nodes[-1,]
```
```{r}
library(stringi)
```

## create files for pathway analysis
```{r}
for (i in 1:nrow(table_nodes)) {
  for (com in 1:8){
    for (type in c("lipidome", "microbiome", "metabolome")) {
      if(table_nodes$Community[i] == com & table_nodes$Type[i] == type) {
        write.table(table_nodes$Metabolites_names[i], paste0("results/microbiome_project/GSEA/", substr(type,1,3),"_", com, ".txt"), row.names = FALSE, quote = FALSE)
      }
    }
  }
}
```

