---
title: "Processing blood samples"
output: html_notebook
---

input metabolon file for lipidomics and metabolomics
1) Normalization (metabolome/lipidome)
2) Data cleaning : remove bottom 15%, remove bottom 15% least variant (metabolome/lipidome)

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/3_layers_COCOMO/")) 
```

```{r}
name_ana <- "microbiome_project"
```

```{r}
library(xlsx)
library(NormalyzerDE)
```

# load data
## clinical data
```{r}
clinical_comp <- read.csv("data/clinical_comp.csv")
clinical <- read.xlsx("data/data_base_200_patients_COCOMO_filtered.xlsx", 1)
```

```{r}
clinical <- merge(clinical, clinical_comp, by = "id")
rownames(clinical) <- clinical$id
clinical_interest <- read.csv("data/clinical_data_of_interest_microbiome_project.csv")
clinical_2 <- clinical_interest$name_table[!clinical_interest$name_table %in% colnames(clinical)]
clinical <- clinical[,colnames(clinical) %in% clinical_interest$name_table]
clinical$id <- rownames(clinical)
write.csv(clinical, paste0("processing/clinical_data_", name_ana, ".csv"))
```

## metabolome
```{r}
data_met <- read.csv("data/cocomo_norm_data_filt.csv")
data_met <- data_met[data_met$SUPER.PATHWAY !="",]
data_met <- data_met[,c(1,13:232)]
names(data_met)[-1] <- gsub("X", "", names(data_met)[-1])
write.csv(clinical, paste0("processing/metabolomics_no_norm_", name_ana, ".csv"))
```

## lipidome
NB : remove total lipids ??
```{r}
data_lip <- read.csv("data/clp_lipidomics_data.csv")
data_lip <- data_lip[data_lip$SUPER_PATHWAY !="",]
data_lip <- data_lip[,c(1, 9:228)]
names(data_lip) <- gsub("X", "", names(data_lip))
write.csv(clinical, paste0("processing/lipidomics_no_norm_", name_ana, ".csv"))
```
## merge everything
```{r}
met <- data_met$BIOCHEMICAL
lip <- data_lip$BIOCHEMICAL
```

```{r}
data_lip <- data.frame(t(data_lip[,-1]))
data_met <- data.frame(t(data_met[,-1]))

colnames(data_lip) <- lip
colnames(data_met) <- met

data_lip$id <- rownames(data_lip)
data_met$id <- rownames(data_met)

data <- merge(data_lip, data_met, by = "id")
data <- merge(clinical, data, by = "id")

write.csv(data, paste0("processing/lipidomics_metabolomics_clinical_no_norm_", name_ana, ".csv"))
```

```{r}
rownames(data) <- data$id
clinical <- data[,2:36]
data <- data[,37:1890]

write.csv(clinical, paste0("processing/clinical_ordered_", name_ana, ".csv"))
```

# Normalization
```{r}
data_2 <- log2(data)
```


```{r}
data_3 <-as.matrix(data_2)
hist(data_3 , prob=T)
curve(dnorm(x,mean(data_3),sd(data_3)),add=T,col="red")
```
## quantile normalization
```{r}
library(preprocessCore)
```

```{r}
data_4 <- normalize.quantiles(as.matrix(data))
```

```{r}
library("car")
qqPlot(data_4)
```

```{r}
data_3 <-as.matrix(data_4)
hist(data_3 , prob=T)
curve(dnorm(x,mean(data_3),sd(data_3)),add=T,col="red")
```
# filtering
## remove 15 % bottom values
```{r}
#n <- data.frame(t(data_2))
#n2 <- data.frame(rowSums(n))
#n3 <- quantile(n2$rowSums.n.,0.15)
#n2 <- rownames(n2)[n2$rowSums.n. > n3]
```
```{r}
#n <- n[rownames(n) %in% n2,]
#met <- rownames(n)
```

## remove low variance values
```{r}
n <- data_2
met <- rownames(n)

n4 <- n[, sapply(n, var) > 0]
#colnames(n) <- met
```

```{r}
rownames(n4) <- gsub("X", "",rownames(n4))
write.csv(n4, paste0("processing/lipidomics_metabolomics_log2_norm_filt", name_ana, ".csv"))
```

```{r}
length(which(colnames(n4) %in% colnames(data_lip)))
length(which(colnames(n4) %in% colnames(data_met)))
```


