---
title: "MOFA pathway analysis"
output: html_notebook
---



### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/3_layers_integration/")) 
```

```{r}
library(xlsx)
library(dplyr)
library(circlize)
library(MOFA2)
library(compositions)
library(stringi)
library(ggalluvial)
```

c()
```{r}
conv_met <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/biochemical_super_path.csv")
names(conv_met)[1] <- "met"
conv_met <- conv_met[conv_met$SUPER.PATHWAY != "",]

conv_lip <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipids_superpathways.csv")
names(conv_lip)[1] <- "item"
tax_table <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/tax_table_COCOMO.csv", header = FALSE)
names(tax_table) <- c("X", "Super_group", "Division", "Class", "Order", "Family", "Genus")
tax_table <- select(tax_table, X, Division, Family, Genus)
names(tax_table)[1] <- "item"
```

```{r}
met_g <- table(conv_met)
lip_g <- table(conv_lip)
mic_g <- table(tax_table$item, tax_table$Division)

met_g2 <- as.data.frame.matrix(table(conv_met))
lip_g2 <- as.data.frame.matrix(table(conv_lip))
mic_g2 <- as.data.frame.matrix(table(tax_table$item, tax_table$Division))
```

```{r}
library(xlsx)
library(dplyr)
library(SNFtool)
library(ROCR)
library(cvAUC)
library(circlize)
library(MOFA2)
library("rhdf5")
library(ComplexHeatmap)
library(ggplot2)
```
```{r}
col <- c("#6b8150", "#2273C3","#EFC144","#868686")
#col_border <- darken(col, 0.5)
```

```{r}
cluster <- read.csv("processing/cluster_SNF_3_omics_3_clusters_plusHC.csv")
cluster <- cluster[!is.na(cluster$cluster),]
cluster$X <- NULL
names(cluster)[2] <- "X"
```

## load clinical
```{r}
clinical <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/clinical_data_microbiome_project.csv")
clinical$Combined_therapy <- gsub("\\/.*", "", clinical$Combined_therapy)
clinical$Combined_therapy <- ifelse(clinical$Combined_therapy %in% c("ABC", "AZT", "TAF", "TDF"), clinical$Combined_therapy, "Others")
clinical_data <- merge(cluster, clinical, by = "X")
clinical_data$CD4_CD8 <- NA
clinical_data$CD4_CD8 <- clinical_data$CD4/clinical_data$CD8
clinical_data <- select(clinical_data, X, cluster, hypertension, Transm_mode, VAT, SAT, CD4_CD8, METS, AGE, BMI, Combined_therapy, X3rd_Drug, GENDER)
names(clinical_data)[1] <- "sample"
clinical_data$group <- "group1"
clinical_data$cluster <- paste0("C", clinical_data$cluster)

```

## Load data
```{r}
data_lip <- read.csv("data/clp_lipidomics_data.csv")
data_met <- read.csv("data/cocomo_norm_data_filt.csv")
data_met <- data_met[data_met$SUPER.PATHWAY !="",]
```

## lipidome
```{r}
data_lip <- data_lip[data_lip$SUPER_PATHWAY !="",]
data_lip <- data_lip[,c(1, 9:228)]
names(data_lip) <- gsub("X", "", names(data_lip))
lip <- data_lip$BIOCHEMICAL
data_lip <- data.frame(t(data_lip[,-1]))
colnames(data_lip) <- lip
grep("Total", colnames(data_lip))
data_lip <- data_lip[rownames(data_lip) %in% clinical_data$sample, ]
data_lip <- data_lip[,1:963]
grep("Total", colnames(data_lip))
data_lip <- log2(data_lip)
#data_lip <- data.frame(t(scale(t(data_lip), scale = TRUE)))
#data_lip <- data_lip[, sapply(data_lip, var) > 0.05]
```

## metabolome
```{r}
data_met <- data_met[,c(1,13:232)]
names(data_met)[-1] <- gsub("X", "", names(data_met)[-1])
met <- data_met$BIOCHEMICAL
data_met <- data.frame(t(data_met[,-1]))
colnames(data_met) <- met
data_met <- data_met[rownames(data_met) %in% clinical_data$sample, ]
data_met <- log2(data_met)
#data_met <- data.frame(t(scale(t(data_met), scale = TRUE)))
#data_met <- data_met[, sapply(data_met, var) > 0.05]
data_met <- data_met[order(match(rownames(data_met), rownames(data_lip))), ]
```


## microbiome
```{r}
data_mi <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/data_count_microbiome.csv")
rownames(data_mi) <- data_mi$X
data_mi$X <- NULL
rownames(data_mi) <- gsub("X","",rownames(data_mi))
data_mi <- data_mi[rownames(data_mi) %in% rownames(data_lip), ]
data_mi <- data.frame(clr(data_mi))
#data_mi <- data.frame(t(scale(t(data_mi), scale = FALSE)))
data_mi <- data_mi[, sapply(data_mi, var) > 0.2]
data_mi <- data_mi[order(match(rownames(data_mi), rownames(data_lip))), ]
```


```{r}
convert_met <- data.frame(met = colnames(data_met), nb = paste0("Met", 1:ncol(data_met)))
colnames(data_met) <- convert_met$nb
```

```{r}
met_g2$met <- rownames(met_g2)
met_g2 <- merge(convert_met, met_g2, by = "met")
rownames(met_g2) <- met_g2$nb
met_g2$nb <- NULL
met_g2$met <- NULL

met_g <- as.matrix(met_g2)
```

## load model (made using defaults parameters)
```{r}
MOFAobject <- load_model("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/mofa_model_4")
MOFAobject
```

```{r}
features_names(MOFAobject)[["Microbiome"]]
```

## add metadata
```{r}
# Sanity check
stopifnot(all(sort(clinical_data$sample)==sort(unlist(samples_names(MOFAobject)))))

# Add sample metadata to the model
samples_metadata(MOFAobject) <- clinical_data
```

```{r}
mic_g <- as.matrix(mic_g2)
lip_g <- as.matrix(lip_g2)
```

```{r}
list_view <- c("Lipids", "Metabolites", "Microbiome")
list_annotation <- list(lip_g, met_g, mic_g)
list_sign <- c("positive", "negative", "all")
list_test <- c("parametric", "cor.adj.parametric", "permutation")
```

```{r}
table_GSEA <- data.frame(View = NA, test = NA, sign = NA, Factor = NA, Pathways = NA)
for (view in 1:length(list_view)) {
  for (test in list_test) {
    for (sign in list_sign) {
      view_sel <- list_view[[view]]
      anno_sel <- list_annotation[[view]]
      enrichment.parametric <- run_enrichment(MOFAobject,
                                              view = view_sel, factors = 1:15,
                                              feature.sets = t(anno_sel),
                                              sign = sign,
                                              statistical.test = test,
                                              p.adj.method = "fdr",
                                              alpha = 0.1,
                                              verbose = FALSE)
      
      m <- enrichment.parametric$sigPathways
      m <- lapply(m, function(x) if(identical(x, character(0))) "none" else x)
      for (i in 1:length(m)) {
        if (length(m[[i]]) > 1) {
          x <- 1:length(m[[i]])
          m[[i]] <- stri_paste(unlist(m[[i]]), collapse='_')
        }
      }
      table_res <- data.frame(View = rep(view_sel, 15), test = rep(test, 15), sign = rep(sign, 15), Factor = 1:15, Pathways = unlist(m))
      table_GSEA <- rbind(table_res, table_GSEA)
    }
  }
}

write.csv(table_GSEA, "processing/table_GSEA_MOFA_factors.csv")
table_GSEA <- table_GSEA[table_GSEA$Pathways != "none",]
table_GSEA <- table_GSEA[!is.na(table_GSEA$sign),]

table_GSEA_1 <- table_GSEA[table_GSEA$test == "parametric",]
table_GSEA_2 <- table_GSEA[table_GSEA$test == "permutation",]
table_GSEA_3 <- table_GSEA[table_GSEA$test == "cor.adj.parametric",]
```
```{r}
table_GSEA_2 <- table_GSEA_2[table_GSEA_2$sign != "all",]
table_GSEA_3 <- table_GSEA_3[table_GSEA_3$sign != "all",]
```

```{r}
table(table_GSEA_1$Factor, table_GSEA_1$Pathways)
```

```{r}
table_GSEA_1 <- table_GSEA_1[table_GSEA_1$sign != "all",]
table_GSEA_1$test <- NULL
table_GSEA_1$Factor <- paste0("Factor", table_GSEA_1$Factor)

table_GSEA <- data.frame(View = NA, sign = NA, Factor = NA, Pathways = NA)
for (i in 1:nrow(table_GSEA_1)){
  x <- paste0(table_GSEA_1$Pathways[i], collapse=" ")
  y <- as.list(strsplit(x, "_")[[1]])
  table_2 <- table_GSEA_1[i,]
  table_2 <- table_2[rep(seq_len(nrow(table_2)), each = length(y)),]
  table_2$Pathways <- y[[1]]
  table_GSEA <- rbind(table_GSEA, table_2)
}

write.csv(table_GSEA, "processing/table_pathway_analysis_MOFA_Factor_2.csv")
```

```{r}
table(table_GSEA$Factor, table_GSEA$Pathways)
```

```{r}
correlations <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/MOFA_correlation_clinical_parameters.csv")
data <- merge(table_GSEA, correlations, by = "Factor")
data$X <- NULL
data$Freq <- 1

data <- data[order(data$Factor),]
data$Factor <- factor(data$Factor, levels = paste0("Factor", 1:15))
data <- data[order(data$View),]
data$Pathways <- factor(data$Pathways, levels = unique(data$Pathways))
data$parameter <- factor(data$parameter, levels = c("cluster", "GENDER", "AGE", "BMI", "VAT", "SAT", "METS", "hypertension", "Transm_mode"))
```

ggplot(as.data.frame(UCBAdmissions),
       aes(y = Freq, axis1 = Gender, axis2 = Dept)) +
  geom_alluvium(aes(fill = Admit), width = 1/12) +
  geom_stratum(width = 1/12, fill = "black", color = "grey") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Gender", "Dept"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle("UC Berkeley admissions and rejections, by sex and department")
  
```{r}
data_pathways <- select(data, View, Factor, Pathways, Freq)
data_clinical <- select(data, View, Factor, parameter, Freq)
```

```{r}
ggplot(as.data.frame(data_clinical),
       aes(y = Freq,
           axis1 = Factor, axis2 = parameter, fill = Factor)) +
  geom_alluvium(aes(fill = View), width = 1/8) +
  geom_stratum(width = 1/8, alpha = .5) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Factor", "Clinical parameters"), expand = c(.05, .05)) 

ggsave("results/figures/sankey_plot_clinical_factors.pdf", width = 6, height = 5)
```

```{r}
ggplot(as.data.frame(data_pathways),
       aes(y = Freq,
           axis1 = Factor, axis2 = Pathways, fill = View)) +
  geom_alluvium(aes(fill = View), width = 1/8) +
  geom_stratum(width = 1/8, alpha = .5) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Factor", "Pathway/Division"), expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1")

ggsave("results/figures/sankey_plot_pathway_analysis_factors.pdf", width = 6, height = 5)
```

```{r}
ggplot(as.data.frame(data),
       aes(y = Freq,
           axis1 = Pathways, axis2 = Factor, axis3 = parameter)) +
  geom_alluvium(aes(fill = View),
                width = 0, knot.pos = 0, reverse = FALSE) +
  guides(fill = FALSE) +
  geom_stratum(width = 1/8, reverse = FALSE) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  scale_x_continuous(breaks = 1:3, labels = c("Pathway/Division", "Factor", "Clinical parameter")) +
  coord_flip()

ggsave("results/figures/pathway_analysis_per_factor_2.pdf", width = 20, height = 6)
```

```{r}
data_2 <- data[,c(2, 4,5,6)]
data_2 <- data_2[!duplicated(data_2[,1:4]),] 
```

```{r}

```


## circos

```{r}
library(GOplot)
```

```{r}
pathways$Genes <- gsub(",", ".", pathways$Genes)
pathways$Genes <- gsub(";", ",", pathways$Genes)
```

```{r}
data_2 <- aggregate(parameter ~ Pathways, data_2, paste, collapse = ", ")
```


ribbon.col = c("#006f62","#62006f", "#6f6200")
```{r}
m <- data.frame(unclass(table(data$parameter, data$Pathways)))
m[m != 0] <- 1                
m$logFC <- runif(9, -2, 2)



pdf("results/figures/chordplot_MOFA.pdf", height = 12, width = 10)
GOChord(m, space = 0.02, gene.space = 0.1, gene.size = 2, border.size = 0.2)
dev.off()
```


```{r}

```

## get weights
```{r}
weights <- get_weights(MOFAobject, 
  views = "all", 
  factors = "all", 
  as.data.frame = TRUE 
)
head(weights)
```

```{r}
names(conv_lip) <- c("feature", "Pathways")
data_lip <- merge(weights, conv_lip, by = "feature")
data_lip$sign <- ifelse(data_lip$value >= 0, "positive", "negative")
```

```{r}
conv_met <- merge(convert_met, conv_met, by = "met")
conv_met$met <- NULL
names(conv_met) <- c("feature", "Pathways")

data_met <- merge(weights, conv_met, by = "feature")
data_met$sign <- ifelse(data_met$value >= 0, "positive", "negative")
```

```{r}
tax_table <- select(tax_table, item, Division)
names(tax_table) <- c("feature", "Pathways")
data_mic <- merge(weights, tax_table, by = "feature")
data_mic$sign <- ifelse(data_mic$value >= 0, "positive", "negative")
```

```{r}
data_lip_2 <- aggregate(data_lip$value, by=list(Factor = data_lip$factor, Pathways = data_lip$Pathways, sign = data_lip$sign), FUN=sum)
data_met_2 <- aggregate(data_met$value, by=list(Factor = data_met$factor, Pathways = data_met$Pathways, sign = data_met$sign), FUN=sum)
data_mic_2 <- aggregate(data_mic$value, by=list(Factor = data_mic$factor, Pathways = data_mic$Pathways, sign = data_mic$sign), FUN=sum)
```

## make individual link
```{r}
for (i in unique(data$parameter)){
  name <- i
  data_2 <- data[data$parameter == i,]
  data_2$weigth <- NULL
  data_3 <- merge(data_2, data_lip_2, by = c("Factor", "Pathways", "sign"))
  data_4 <- merge(data_2, data_met_2, by = c("Factor", "Pathways", "sign"))
  data_5 <- merge(data_2, data_mic_2, by = c("Factor", "Pathways", "sign"))
  data_6 <- rbind(data_3, data_4, data_5)
  ggplot(data=data_6, aes(y=abs(x), x=Factor, fill=Pathways)) +
    geom_bar(stat="identity")+
    scale_fill_brewer(palette="Paired")+
    theme_minimal()+xlab(name)+ylab("absolute weight")
  ggsave(paste0("results/figures/barplots/pathway_analysis_per_factor_3_", i,".pdf"), width = 4 + nrow(data_2))
  
  ggplot(as.data.frame(data_2),
       aes(y = Freq,
           axis1 = Pathways, axis2 = Factor, axis3 = parameter)) +
  geom_alluvium(aes(fill = View),
                width = 0, knot.pos = 0, reverse = FALSE) +
  guides(fill = FALSE) +
  geom_stratum(width = 1/8, reverse = FALSE) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  scale_x_continuous(breaks = 1:3, labels = c("Pathway/Division", "Factor", "Clinical parameter")) +
  coord_flip()
  ggsave(paste0("results/figures/sankeys/pathway_analysis_per_factor_3_", i,".pdf"), width = 4 + nrow(data_2))

}
```

