---
title: "R Notebook"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/3_layers_integration/")) 
```


/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG2.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG1.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G1G2.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_CTG2.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G2G3.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G1G3.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G1G2.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG3.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G3G2.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G1G3.txt
/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_CTG3.txt


## load pathways metabolome
```{r}
cutoff <- 0.2

c1 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_CTG1.txt")
c1 <- c1[c1$Adjusted.P.value < cutoff,]
c2 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_CTG2.txt")
c2 <- c2[c2$Adjusted.P.value < cutoff,]
c3 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_CTG3.txt")
c3 <- c3[c3$Adjusted.P.value < cutoff,]
c4 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G1G2.txt")
c4 <- c4[c4$Adjusted.P.value < cutoff,]
c5 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G1G3.txt")
c5 <- c5[c5$Adjusted.P.value < cutoff,]
c6 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/met_G3G2.txt")
c6 <- c6[c6$Adjusted.P.value < cutoff,]




c1$Comp <- "CTG1"
c2$Comp <- "CTG2"
#c3$Comp <- "CTG3"

c4$Comp <- "G1G2"
c5$Comp <- "G1G3"
c6$Comp <- "G2G3"
```

```{r}
msea <- rbind(c1, c2, c4, c5, c6)
msea <- msea[msea$Overlap != "1/1",]
```

```{r}
msea$value <- 1
msea$Nb_metabolites <- as.numeric(gsub("\\/.*", "", msea$Overlap))
```
```{r}
library(Cairo)
library(viridis)
```

```{r}
cairo_pdf("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/figures/3_clusters_gseapy_0_2_metabolites.pdf", height = 4, width = 4,family="mono")
 a <- ggplot(msea, aes(x = value, y = Term, size = Nb_metabolites))+
        geom_point()+ scale_x_continuous(limits=c(0.9999999999, 1.0000000001))+ theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank(), axis.text.y = element_text(size=8, color = "black"))+ scale_color_viridis(discrete = TRUE)+ facet_grid(~ Comp)
a
dev.off()
#ggsave("results/figures/Figure1e_MSEA_gseapy_Metabolon.pdf", height = 4, width = 12)
```
```{r}
cutoff <- 0.2

c1 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG1.txt")
c1 <- c1[c1$Adjusted.P.value < cutoff,]
c2 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG2.txt")
c2 <- c2[c2$Adjusted.P.value < cutoff,]
c3 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_CTG3.txt")
c3 <- c3[c3$Adjusted.P.value < cutoff,]
c4 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G1G2.txt")
c4 <- c4[c4$Adjusted.P.value < cutoff,]
c5 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G1G3.txt")
c5 <- c5[c5$Adjusted.P.value < cutoff,]
c6 <- read.delim("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/MSEA/lip_G2G3.txt")
c6 <- c6[c6$Adjusted.P.value < cutoff,]




c1$Comp <- "CTG1"
c2$Comp <- "CTG2"
c3$Comp <- "CTG3"

c4$Comp <- "G1G2"
c5$Comp <- "G1G3"
c6$Comp <- "G2G3"
```

```{r}
msea <- rbind(c1, c2, c3, c4, c5, c6)
msea <- msea[msea$Overlap != "1/1",]
```

```{r}
msea$value <- 1
msea$Nb_lipids <- as.numeric(gsub("\\/.*", "", msea$Overlap))
```
```{r}
library(Cairo)
library(viridis)
```

```{r}
cairo_pdf("/home/flomik/Desktop/Code-PHD/3_layers_integration/results/figures/3_clusters_gseapy_0_2_lipids.pdf", height = 4, width = 8,family="mono")
 a <- ggplot(msea, aes(x = value, y = Term, size = Nb_lipids))+
        geom_point()+ scale_x_continuous(limits=c(0.9999999999, 1.0000000001))+ theme(axis.line=element_blank(),axis.text.x=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank(), axis.text.y = element_text(size=8, color = "black"))+ scale_color_viridis(discrete = TRUE)+ facet_grid(~ Comp)
a
dev.off()
#ggsave("results/figures/Figure1e_MSEA_gseapy_Metabolon.pdf", height = 4, width = 12)
```
```{r}

```

