---
title: "R Notebook"
output: html_notebook
---

### enter your path to folder
```{r}
path <- "/home/flomik/Desktop/Code-PHD/Git/HIV_multiomics"
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath(path)) 
```


```{r}
library(xlsx)
library(gplots)
library(ComplexHeatmap)
library(circlize)
library(limma)
library("colorspace")
library(ggplot2)
library(dplyr)
library(scales)
```

```{r}
col <- c("#6b8150", "#2273C3","#EFC144","#868686")
col_border <- darken(col, 0.5)
```

```{r}
data_met <- read.csv("processing/metabolomics_log2_plusHC.csv", check.names = FALSE)
data_met$cluster <- factor(data_met$cluster, levels = c("Ctrl", 1, 2, 3))
cluster <- data_met[,c(1,2)]
data_lip <- read.csv("processing/lipidomics_log2_plusHC.csv", check.names = FALSE, row.names = NULL)
names(data_lip)[1] <- "X"
data_lip <- merge(cluster, data_lip, by = "X")
data_lip$cluster <- factor(data_lip$cluster, levels = c("Ctrl", 1, 2, 3))
```

```{r}
clinical <- read.csv("processing/clinical_data_microbiome_project.csv")
clinical <- clinical[,c(1, 8, 9)]
```

```{r}
mic <- read.csv("processing/relative_abundance_table.csv", row.names = 1)
mic <- data.frame(t(mic))
mic$X <- gsub("X","", rownames(mic))
```

```{r}
c1 <- data_met[,c(1, 2)]
c2 <- data_lip[,c(1, 2)]
```

## LIMMA metabolites
```{r}
pca_data_2 <- merge(clinical, data_met, by = "X", all.y = TRUE)

pca_data_2 <- pca_data_2[pca_data_2$cluster != "Ctrl",]
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group + cd4 + trans)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(
                             C3_C1 = comp[1],
                             C2_C1 = comp[2],
                             C2_C3 = comp[3],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```

```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.1], paste0("processing/GSEA/met_", comp[i],"_corrected.txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_metabolites_with_corrected.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.1, ]

path_results <- paste0("results/LIMMA_results_metabolites_with_filt_corrected.csv")

write.table(DF_top_table, file = path_results)
```
## pca metabolomics
```{r}
pca_data <- pca_data_2[,colnames(pca_data_2) %in% c("cluster", DF_top_table$BIOCHEMICAL)]
pca_data$X <- NULL
pca_data[is.na(pca_data)] <-  0
#remove <- which(apply(pca_data, 2, var)==0)
#pca_data <- pca_data[,-remove]
pca <- prcomp(pca_data[,-1], scale. =  FALSE)
df_out <- as.data.frame(pca$x)
df_out$group <- as.factor(pca_data$cluster)
head(df_out)
```

```{r}
x <- project.pca.proportionvariances <- ((pca$sdev^2) / (sum(pca$sdev^2)))*100
a1 <- x[1]
a2 <- x[2]
barplot(project.pca.proportionvariances, cex.names=1, xlab=paste("Principal component (PC), 1-", length(pca$sdev)), ylab="Proportion of variation (%)", main="Scree plot", ylim=c(0,100))
PCAloadings <- data.frame(Variables = rownames(pca$rotation), pca$rotation)
PCAloadings_2 <- PCAloadings
x <- abs(PCAloadings_2$PC1) < 0.1 & abs(PCAloadings_2$PC2) < 0.1
PCAloadings_2$PC1[x] <- NA
PCAloadings_2$PC2[x] <- NA
PCAloadings_2$Variables[x] <- NA
```
# build ggplot dataframe with points (x,y) and corresponding groups (cluster)

# calculate group centroid locations
centroids <- aggregate(cbind(x,y)~cluster,data=gg,mean)
# merge centroid locations into ggplot dataframe
gg <- merge(gg,centroids,by="cluster",suffixes=c("",".centroid"))
# generate star plot...
ggplot(gg) +
  geom_point(aes(x=x,y=y,color=cluster), size=3) +
  geom_point(data=centroids, aes(x=x, y=y, color=cluster), size=4) +
  geom_segment(aes(x=x.centroid, y=y.centroid, xend=x, yend=y, color=cluster))

```{r}
df_out$group <- factor(df_out$group, levels = c(1, 2, 3))
```
```{r}
col <- c("#2273C3","#EFC144","#868686")
```

```{r}
gg <- data.frame(cluster=factor(df_out$group), x=df_out$PC1, y=df_out$PC2)
centroids <- aggregate(cbind(x,y)~cluster,data=gg,mean)
gg <- merge(gg,centroids,by="cluster",suffixes=c("",".centroid"))

ggplot(gg)+
  geom_point(aes(x=x,y=y,color=cluster, fill = cluster), size=4) + 
  geom_segment(aes(x=x.centroid, y=y.centroid, xend=x, yend=y, color=cluster))+
  geom_point(data=centroids, aes(x=x, y=y, color=cluster, fill = cluster), size = 5, alpha = 1, shape = 19)+
  stat_ellipse(geom = "polygon", aes(x=x,y=y,color=cluster, fill = cluster), alpha = 0.1)+
  theme(legend.title=element_text(size=15),legend.text=element_text(size=12),
        legend.key.size=unit(0.7,"line"),plot.title = element_text(hjust = 0.5),
        axis.title.y=element_text(size=15),axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=18),axis.text.x=element_text(size=18))+ scale_color_manual(values=col)+ scale_fill_manual(values=col)+
  xlab(paste0("PC1: ",round(a1,0),"% variance"))+
  ylab(paste0("PC2: ",round(a2,0),"% variance"))+ 
    theme(axis.title.x = element_text(size=16, face = "bold",color="black"),
          axis.title.y = element_text(size=16, face = "bold",color="black"),
          axis.text = element_text(color="black",size=16, face = "bold"),
          axis.text.x =element_text(color="black",size=16, face = "bold"),
          panel.border = element_blank(),
          axis.ticks.x = element_blank(),
          strip.text.x = element_text(size = 9,color="black"),
          panel.background = element_rect(fill = "white", colour = "white", size = 0.5, linetype = "solid"),
          panel.grid.major = element_line(size = 0.5, linetype ='solid', colour = "grey"),
          panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
          plot.title = element_text(hjust = 0.5, size = 20, face = "bold", vjust=2),
          legend.background=element_rect(fill = alpha("white", 0)))

path_fig <- paste0("results/figures/PCA_clustering_97_patients_3_clusters_metabolome_test_corrected.pdf")
dev.copy(pdf, path_fig, height = 4, width = 6)
dev.off()
```


## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_metabolome_corrected", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## LIMMA lipids
```{r}
pca_data_2 <- merge(clinical, data_lip, by = "X", all.y = TRUE)

pca_data_2 <- pca_data_2[pca_data_2$cluster != "Ctrl",]
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group + cd4 + trans)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(
                             C3_C1 = comp[1],
                             C2_C1 = comp[2],
                             C2_C3 = comp[3],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```
```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.05], paste0("processing/GSEA/lip_corrected_", comp[i],".txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_lipids_with_corrected.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.05, ]

path_results <- paste0("results/LIMMA_results_lipids_with_filt_corrected.csv")

write.table(DF_top_table, file = path_results)

table(DF_top_table$Comp)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lipidomics_corrected", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## LIMMA metabolites
```{r}
pca_data_2 <- merge(clinical, data_met, by = "X", all.y = TRUE)

pca_data_2 <- pca_data_2[pca_data_2$cluster != "Ctrl",]
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(
                             C3_C1 = comp[1],
                             C2_C1 = comp[2],
                             C2_C3 = comp[3],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```

```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.1], paste0("processing/GSEA/met_", comp[i],".txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_metabolites_with.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.1, ]

path_results <- paste0("results/LIMMA_results_metabolites_with_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_metabolome", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## LIMMA lipids
```{r}
pca_data_2 <- merge(clinical, data_lip, by = "X", all.y = TRUE)

pca_data_2 <- pca_data_2[pca_data_2$cluster != "Ctrl",]
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(
                             C3_C1 = comp[1],
                             C2_C1 = comp[2],
                             C2_C3 = comp[3],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```
```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.05], paste0("processing/GSEA/lip_", comp[i],".txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_lipids_with.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.05, ]

path_results <- paste0("results/LIMMA_results_lipids_with_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lipidomics", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```



## LIMMA metabolites
```{r}
pca_data_2 <- merge(clinical, data_met, by = "X", all.y = TRUE)
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group1-groupCtrl", "group2-groupCtrl", "group3-groupCtrl", "group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(Ctrl_C1 = comp[1],
                             Ctrl_C2 = comp[2],
                             Ctrl_C3 = comp[3],
                             C3_C1 = comp[4],
                             C2_C1 = comp[5],
                             C2_C3 = comp[6],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```


```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.1], paste0("processing/GSEA/met_", comp[i],"_HC.txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_metabolites_with_HC.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.1, ]

path_results <- paste0("results/LIMMA_results_metabolites_with_HC_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_metabolome_HC_", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## LIMMA lipids
```{r}
pca_data_2 <- merge(clinical, data_lip, by = "X", all.y = TRUE)
```

```{r}
group <- factor(pca_data_2$cluster, levels = unique(pca_data_2$cluster))
pca_data_2$Transm_mode[is.na(pca_data_2$Transm_mode) | pca_data_2$Transm_mode  == 7] <- "8"
trans <- as.factor(pca_data_2$Transm_mode)
cd4 <- pca_data_2$CD4
design <- model.matrix( ~ 0 + group)
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1:4)]), design)
```

```{r}
comp <- c("group1-groupCtrl", "group2-groupCtrl", "group3-groupCtrl", "group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(Ctrl_C1 = comp[1],
                             Ctrl_C2 = comp[2],
                             Ctrl_C3 = comp[3],
                             C3_C1 = comp[4],
                             C2_C1 = comp[5],
                             C2_C3 = comp[6],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```

```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.05], paste0("processing/GSEA/lip_", comp[i],"_HC.txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_lipids_with_HC.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.05, ]

path_results <- paste0("results/LIMMA_results_lipids_with_HC_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lipidomics_HC", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```







## check overlap
```{r}
c1 <- read.csv("results/LIMMA_results_lipids_with_HC_filt.csv", sep = "")
c2 <- read.csv("results/LIMMA_results_lipids_with_filt.csv", sep = "")
c3 <- read.csv("results/LIMMA_results_lipids_with_filt_corrected.csv", sep = "")

c4 <- read.csv("results/LIMMA_results_metabolites_with_HC_filt.csv", sep = "")
c5 <- read.csv("results/LIMMA_results_metabolites_with_filt.csv", sep = "")
c6 <- read.csv("results/LIMMA_results_metabolites_with_filt_corrected.csv", sep = "")
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(c1$BIOCHEMICAL[c1$Comp == "group2-group1"])
b <- list(c2$BIOCHEMICAL[c2$Comp == "group2-group1"])
c <- list(c3$BIOCHEMICAL[c3$Comp == "group2-group1"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_met_C1_C2", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)

a <- list(c1$BIOCHEMICAL[c1$Comp == "group3-group1"])
b <- list(c2$BIOCHEMICAL[c2$Comp == "group3-group1"])
c <- list(c3$BIOCHEMICAL[c3$Comp == "group3-group1"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_met_C1_C3", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)

a <- list(c1$BIOCHEMICAL[c1$Comp == "group3-group2"])
b <- list(c2$BIOCHEMICAL[c2$Comp == "group3-group2"])
c <- list(c3$BIOCHEMICAL[c3$Comp == "group3-group2"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_met_C2_C3", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(c4$BIOCHEMICAL[c4$Comp == "group2-group1"])
b <- list(c5$BIOCHEMICAL[c5$Comp == "group2-group1"])
c <- list(c6$BIOCHEMICAL[c6$Comp == "group2-group1"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lip_C1_C2", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)

a <- list(c4$BIOCHEMICAL[c4$Comp == "group3-group1"])
b <- list(c5$BIOCHEMICAL[c5$Comp == "group3-group1"])
c <- list(c6$BIOCHEMICAL[c6$Comp == "group3-group1"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lip_C1_C3", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)

a <- list(c4$BIOCHEMICAL[c4$Comp == "group3-group2"])
b <- list(c5$BIOCHEMICAL[c5$Comp == "group3-group2"])
c <- list(c6$BIOCHEMICAL[c6$Comp == "group3-group2"])

venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lip_C2_C3", category.names = c("Model_HC" , "Model" , " Model_corrected"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```
