---
title: "R Notebook"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/3_layers_integration/")) 
```

```{r}
library(xlsx)
library(gplots)
library(ComplexHeatmap)
library(circlize)
library(limma)
library("colorspace")
library(ggplot2)
library(dplyr)
library(scales)
```

```{r}
col <- c("#6b8150", "#2273C3","#EFC144","#868686")
col_border <- darken(col, 0.5)
```

```{r}
cluster <- read.csv("processing/cluster_SNF_3_omics_3_clusters_plusHC.csv")
cluster$X <- NULL
names(cluster)[2] <- "X"
```

```{r}
data_lip <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipidomics_log2_plusHC.csv", check.names = FALSE, row.names = NULL)
data_lip$cluster <- factor(data_lip$cluster, levels = c("Ctrl", 1, 2, 3))
data_met <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/metabolomics_log2_plusHC.csv", check.names = FALSE)
data_met$cluster <- factor(data_met$cluster, levels = c("Ctrl", 1, 2, 3))
```

```{r}
mic <- read.csv("processing/relative_abundance_table.csv", row.names = 1)
mic <- data.frame(t(mic))
mic$X <- gsub("X","", rownames(mic))
```

```{r}
c1 <- select(data_met, X, cluster)
c2 <- select(data_lip, X, cluster)

clustering <- merge(c1, cluster, by = "X")
clustering <- merge(c2, clustering, by = "X")
```


## LIMMA metabolites
```{r}
pca_data_2 <- data_met
```

```{r}
group <- as.factor(pca_data_2$cluster)
design <- model.matrix( ~ 0 + group)
design
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1,2)]), design)
```

```{r}
comp <- c("group1-groupCtrl", "group2-groupCtrl", "group3-groupCtrl", "group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(Ctrl_C1 = comp[1],
                             Ctrl_C2 = comp[2],
                             Ctrl_C3 = comp[3],
                             C3_C1 = comp[4],
                             C2_C1 = comp[5],
                             C2_C3 = comp[6],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```

```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.1], paste0("processing/GSEA/met_", comp[i],".txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_metabolites_with_HC.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.1, ]

path_results <- paste0("results/LIMMA_results_metabolites_with_HC_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-group1"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group1"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-group2"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_metabolome", category.names = c("SNF-2vsSNF-1" , "SNF-3vsSNF-1" , " SNF-3vsSNF-2"), fill = myCol, euler.d = TRUE,
	cex = 1.5,
	cat.cex = 0.8)
```

## LIMMA lipids
```{r}
pca_data_2 <- data_lip
```

```{r}
group <- as.factor(pca_data_2$cluster)
design <- model.matrix( ~ 0 + group)
design
```

```{r}
fit <- lmFit(t(pca_data_2[,-c(1,2)]), design)
```

```{r}
design
```

```{r}
comp <- c("group1-groupCtrl", "group2-groupCtrl", "group3-groupCtrl", "group3-group1", "group2-group1", "group3-group2")
```

```{r}
cont.matrix <- makeContrasts(Ctrl_C1 = comp[1],
                             Ctrl_C2 = comp[2],
                             Ctrl_C3 = comp[3],
                             C3_C1 = comp[4],
                             C2_C1 = comp[5],
                             C2_C3 = comp[6],
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
```
```{r}
top_table <- data.frame(BIOCHEMICAL = NA , logFC= NA , P.Value= NA , adj.P.Val= NA , Comp= NA )

for (i in 1:length(comp)) {
  top_table_1 <- topTable(fit.cont, coef=i, adjust="BH", n = Inf, sort.by = "P")
  top_table_1$Comp <- comp[i]
  top_table_1$BIOCHEMICAL <- rownames(top_table_1)
  top_table_1 <- select(top_table_1, BIOCHEMICAL, logFC, P.Value, adj.P.Val, Comp)
  write.table(top_table_1$BIOCHEMICAL[top_table_1$adj.P.Val < 0.05], paste0("processing/GSEA/lip_", comp[i],".txt"), quote = FALSE, row.names = F)
  top_table <- rbind(top_table, top_table_1)
}

top_table <- top_table[complete.cases(top_table),]
path_table <- paste0("results/LIMMA_results_lipids_with_HC.csv")

print(path_table)
write.csv(top_table, file =path_table)

DF_top_table <-top_table[top_table$adj.P.Val < 0.05, ]

path_results <- paste0("results/LIMMA_results_lipids_with_HC_filt.csv")

write.table(DF_top_table, file = path_results)
```

## make ven diagram 3 comp
```{r}
library(nVennR)

a <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group1-groupCtrl"])
b <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group2-groupCtrl"])
c <- list(DF_top_table$BIOCHEMICAL[DF_top_table$Comp == "group3-groupCtrl"])
```

```{r}
# Load library
library(VennDiagram)
 # Prepare a palette of 3 colors with R colorbrewer:
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")

# Chart
venn.diagram(x = list(a[[1]], b[[1]], c[[1]]), filename = "results/figures/ven_lipidome", category.names = c("Ctrl vs C1" , "Ctrl vs C2" , " Ctrl vs C3"), fill = myCol)
```
/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipids_pathways.csv
/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipids_superpathways.csv
/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/metabolites_pathways.csv



```{r}
table_item <- read.csv("results/LIMMA_results_metabolites_with_HC.csv")
conv <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/biochemical_super_path.csv")
input_bar <- merge(table_item, conv, by = "BIOCHEMICAL")
input_bar_2 <- select(input_bar, SUPER.PATHWAY, Comp, adj.P.Val)
input_bar_2 <- input_bar_2[input_bar_2$adj.P.Val < 0.1,]

input_bar_2 <- select(input_bar_2, SUPER.PATHWAY, Comp)
input_bar <- select(input_bar, SUPER.PATHWAY, Comp)

d3 <- data.frame(table(input_bar))
names(d3)[c(1,3)] <-c("Category", "count")
d3 <- d3[d3$count != 0,]
d3 <- d3[unique(d3$Category),]
d3$Comp <- "Total"
d3$fraction <- d3$count/sum(d3$count)
d3$N <- sum(d3$count)

d4 <- data.frame(table(input_bar_2))
names(d4)[c(1,3)] <-c("Category", "count")
d4 <- d4[d4$count != 0,]

```

```{r}
d4$fraction <- NA
d4$N <- NA
for (comp in unique(d4$Comp)) {
  d4$fraction[which(d4$Comp == comp)] <- d4$count[which(d4$Comp == comp)]/sum(d4$count[which(d4$Comp == comp)])
  d4$N[which(d4$Comp == comp)] <- sum(d4$count[which(d4$Comp == comp)])
  print(comp)
  
}
```
```{r}
names(d3)
```
```{r}
names(d4)
```
```{r}
d5 <- rbind(d3, d4)
```

```{r}
d5$Group <- paste0(d5$Comp, ", N=",d5$N)
d5$Group <- factor(d5$Group, levels = rev(c(unique(d5$Group)[grep("Total", unique(d5$Group))],
                                        unique(d5$Group)[grep("group1-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group2-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group2-group1", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-group1", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-group2", unique(d5$Group))]
                                        )))
```

```{r}
mycols_met <- c("#34558b", "#f0daa4", "#eaac9d", "#798fa8", "#fd823e", "#117893", "#d13b40", "#ffaf12", "#a2553a")

ggplot(d5, aes(fill=Category, y=fraction, x=Group))+ 
    geom_bar(stat="identity", alpha = 1, width = .7, colour="white", lwd=0.1) +
  labs(x="Comparison",y="Comorbidities frequency")+
    scale_fill_manual(values = rev(mycols_met))+
  geom_text(aes(label=ifelse(fraction >= 0.07, paste0("n = ", count, ", \n", round(fraction*100, 0), "%"),"")), position=position_stack(vjust=0.5), colour="white")+
  coord_flip() +
  scale_y_continuous(labels = percent_format()) +
  labs(y="", x="")

ggsave(paste0("results/figures/barplot_comparisons_LIMMA_Metabolites.pdf"), width = 15, height = 8)
```

```{r}
library(RColorBrewer)
mycols_lip <- c(brewer.pal(name="Dark2", n = 8), brewer.pal(name="Paired", n = 6))
```

/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipids_pathways.csv

/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/metabolites_pathways.csv



```{r}
table_item <- read.csv("results/LIMMA_results_lipids_with_HC.csv")
conv <- read.csv("/home/flomik/Desktop/Code-PHD/3_layers_integration/processing/lipids_superpathways.csv")
input_bar <- merge(table_item, conv, by = "BIOCHEMICAL")
input_bar_2 <- select(input_bar, SUPER_PATHWAY, Comp, adj.P.Val)
input_bar_2 <- input_bar_2[input_bar_2$adj.P.Val < 0.05,]

input_bar_2 <- select(input_bar_2, SUPER_PATHWAY, Comp)
input_bar <- select(input_bar, SUPER_PATHWAY, Comp)

d3 <- data.frame(table(input_bar))
names(d3)[c(1,3)] <-c("Category", "count")
d3 <- d3[d3$count != 0,]
d3 <- d3[unique(d3$Category),]
d3$Comp <- "Total"
d3$fraction <- d3$count/sum(d3$count)
d3$N <- sum(d3$count)

d4 <- data.frame(table(input_bar_2))
names(d4)[c(1,3)] <-c("Category", "count")
d4 <- d4[d3$count != 0,]

```

```{r}
d4$fraction <- NA
d4$N <- NA
for (comp in unique(d4$Comp)) {
  d4$fraction[which(d4$Comp == comp)] <- d4$count[which(d4$Comp == comp)]/sum(d4$count[which(d4$Comp == comp)])
  d4$N[which(d4$Comp == comp)] <- sum(d4$count[which(d4$Comp == comp)])
  print(comp)
  
}
```
```{r}
names(d3)
```
```{r}
names(d4)
```
```{r}
d5 <- rbind(d3, d4)
```

```{r}
d5$Group <- paste0(d5$Comp, ", N=",d5$N)
d5$Group <- factor(d5$Group, levels = rev(c(unique(d5$Group)[grep("Total", unique(d5$Group))],
                                        unique(d5$Group)[grep("group1-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group2-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-groupCtrl", unique(d5$Group))],
                                        unique(d5$Group)[grep("group2-group1", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-group1", unique(d5$Group))],
                                        unique(d5$Group)[grep("group3-group2", unique(d5$Group))]
                                        )))
```

```{r}
mycols_met <- c("#34558b", "#f0daa4", "#eaac9d", "#798fa8", "#fd823e", "#117893", "#d13b40", "#ffaf12", "#a2553a")

ggplot(d5, aes(fill=Category, y=fraction, x=Group))+ 
    geom_bar(stat="identity", alpha = 1, width = .7, colour="white", lwd=0.1) +
  labs(x="Comparison",y="Comorbidities frequency")+
    scale_fill_manual(values = rev(mycols_lip))+
  geom_text(aes(label=ifelse(fraction >= 0.07, paste0("n = ", count, ", \n", round(fraction*100, 0), "%"),"")), position=position_stack(vjust=0.5), colour="white")+
  coord_flip() +
  scale_y_continuous(labels = percent_format()) +
  labs(y="", x="")

ggsave(paste0("results/figures/barplot_comparisons_LIMMA_Lipids.pdf"), width = 15, height = 8)
```

