---
title: "Preprocessing input files"
output: html_notebook
---




### enter your path to folder
```{r}
path <- "/home/flomik/Desktop/Code-PHD/Git/HIV_multiomics"
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath(path)) 
```

```{r}
library(xlsx)
```

/home/flomik/Desktop/Code-PHD/Git/HIV_multiomics/data/data_base_200_patients_COCOMO_filtered.xlsx
/home/flomik/Desktop/Code-PHD/Git/HIV_multiomics/data/Original_Scale_Data_CLP.xlsx
/home/flomik/Desktop/Code-PHD/Git/HIV_multiomics/data/Supplementary Data 1.xlsx

## Lipidomics
```{r}

```

## Metabolomics
```{r}

```

## Microbiome
### select rarefied files for genus, family and phylum
```{r}
data_genus <- read.xlsx("data/COCOMO-request.SHH.withfamilyandphylum.xlsx", 2)
write.csv(data_genus, "data/microbiome_COCOMO_genus.csv")

data_family <- read.xlsx("data/COCOMO-request.SHH.withfamilyandphylum.xlsx", 4)
write.csv(data_family, "data/microbiome_COCOMO_family.csv")

data_phylum <- read.xlsx("data/COCOMO-request.SHH.withfamilyandphylum.xlsx", 6)
write.csv(data_phylum, "data/microbiome_COCOMO_phylum.csv")
```

